Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-16 paper.tex
remove old comment
handle some question from lukas in evaluation

  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G'^i_o)$ 
   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$; \lukas{how does the merge work? do we explain this?}

@@

\section{Evaluation}
\label{sec:eval}

%% Put three subfigures of simulations, start, after, end
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Initial state}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Simulation of the proposed approach. 
  The dots represent the robots, 
  the circle area on the top right represents the phenomenon to be monitored, 
  and the grey area represents the area covered by the robots.}
  \label{fig:simulations}
\end{figure}
\ga{Page budget: 2 pages \\}
%\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
%I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.}
To test the effectiveness of the proposed approach, 
 we decided to experiment with a case study related to swarm robotics, 
 specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
 or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
 \lukas{``Initially we consider a stationary phenomenon before expanding towards a moving phenomenon. Additionally, the phenomenon can change its shape during runtime. Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. Figure~\ref{fig:simulations} illustrates a set of 25 agents and a single, stationary phenomenon with a with a 3D Gaussian distribution.''}

\subsection{Scenario}
The proposed simulation consists 
 of a set of agents placed in a 2D space capable of perceiving the intensity 
 of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
 (e.g., camera, temperature sensor, etc.). 
%
Additionally, each agent has a coverage range $\omega$ 
 that describes the area it can monitor. 
 Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
 which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. \lukas{$\mathbb{O}$  not defined.}
From the neighbourhood, 
 each agent can also perceive the direction of the other agents.
%
Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
 which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
 \lukas{drone, robot, agent, node...}
\subsection{Goal}
 The objective of this scenario is threefold:
\begin{enumerate}
\item Maximise the number of drones within the phenomena
\item Minimise the number of drones without a neighbourhood
\item Maximise the coverage of the system
\end{enumerate}

@@

% Regarding the first point, the reward function is simple and is defined as:
Formally, we define the reward of a drone being within the phenomenon as
 \begin{equation*}
 R^a_{v} = \begin{cases}
  1 & \text{if } \zeta_v > 0 \\
  0 & \text{otherwise.} 
 \end{cases}
 \end{equation*}

 I.e., a drone is considered within the phenomena as soon as the drone can sense the phenomenon.
This will lead the system to prefer a configuration in which every agent is present within the phenomenon.
%
The second element in the objective function ensures cohesion among the agents. 
 This is important because if the system breaks into many scattered agents, 
 the observability of the phenomenon is reduced, 
 limiting the ability of the agents to move appropriately in the environment.
% they are not able to move in the environment due to the reduced observability of the system. 
 In this case, the reward is defined as:
 \begin{equation*}
 R^{\mathcal{N}}_{v} = \begin{cases}
  1 & \text{if } \mathcal{N} > 0 \\
  0 & \text{otherwise} 
 \end{cases}
 \end{equation*}

 Finally, to maximise the coverage, 
  we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar. 
  \lukas{minimum / maximum distance to what? the next neighbour? measuring the phenomena?}
%
This means that the average distance will tend towards the one expressed by the viewing range of each node:
% 
\begin{equation*}
R^{C}_{v} = \frac{d_{max} - d_{min}}{\mathbb{O}}
\end{equation*}
where 
$d_{max}$ and $d_{min}$ describes the maximum and minimum distance of the agent to \lukas{??}, respectively. 
This distance is normalized by the maximum communication range of each agent $\mathbb{O}$.
\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
%

@@

\subsection{Discussion and Results}
%% Subfigures of coverage-in-time and reward-in-time

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Coverage in time}
    \label{fig:coverage}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward in time}
    \label{fig:reward}
  \end{subfigure}
  \caption{Coverage and reward in time}
  \label{fig:coverage-reward}
\end{figure}
%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test

>>>>

  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G'^i_o)$ 
   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$;

@@

\section{Evaluation}
\label{sec:eval}
\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{imgs/start.png}
  \label{fig:example}
  \caption{Simulation of the case study scenario in Alchemist. 
  The dots represent the robots, 
  the circle area on the top right represents the phenomenon to be monitored, 
  and the grey area represents the area covered by the robots.}
\end{figure}
\ga{Page budget: 2 pages \\}
%\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
%I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.}
To test the effectiveness of the proposed approach, 
 we decided to experiment with a case study related to swarm robotics, 
 specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
 or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
 %
 Initially, 
  we perform the training phase using a stationary phenomenon before expanding towards a moving phenomenon in the test phase. 
  %Additionally, the phenomenon can change its shape during runtime. 
  Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. 
  Figure~\ref{fig:example} illustrates a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian distribution.''

\subsection{Scenario}
The proposed simulation consists 
 of a set of agents placed in a 2D space capable of perceiving the intensity 
 of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
 (e.g., camera, temperature sensor, etc.). 
%
Additionally, each agent has a coverage range $\omega$ 
 that describes the area it can monitor. 
 Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
 which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. \lukas{$\mathbb{O}$  not defined.}
From the neighbourhood, 
 each agent can also perceive the direction of the other agents.
%
Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
 which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
 \lukas{drone, robot, agent, node...}
\subsection{Goal}
 The objective of this scenario is threefold:
\begin{enumerate}
\item Maximise the number of agents within the phenomena
\item Minimise the number of agents without a neighbourhood
\item Maximise the coverage of the system
\end{enumerate}

@@

% Regarding the first point, the reward function is simple and is defined as:
Formally, we define the reward of an agent being within the phenomenon as
 \begin{equation*}
 R^a_{v} = \begin{cases}
  1 & \text{if } \zeta_v > 0 \\
  0 & \text{otherwise.} 
 \end{cases}
 \end{equation*}
Namely, an agent is considered within the phenomena as soon as the drone can sense the phenomenon.
%
This will lead the system to prefer a configuration in which every agent is present within the phenomenon.
%
The second element in the objective function ensures cohesion among the agents. 
 This is important because if the system breaks into many scattered agents, 
 the observability of the phenomenon is reduced, 
 limiting the ability of the agents to move appropriately in the environment.
% they are not able to move in the environment due to the reduced observability of the system. 
 In this case, the reward is defined as:
 \begin{equation*}
 R^{\mathcal{N}}_{v} = \begin{cases}
  1 & \text{if } \mathcal{N} > 0 \\
  0 & \text{otherwise} 
 \end{cases}
 \end{equation*}

Finally, to maximise the coverage, 
  we define a reward function that favours the maximum distance between the agents
  equal to the coverage range $\omega$.
%
This means that the average distance will tend towards the one expressed by the viewing range of each node:
% 
\begin{equation*}
R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
\end{equation*}
where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
%
%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
%

@@

\subsection{Discussion and Results}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Initial configuration}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.}
  \label{fig:simulations}
\end{figure}
\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Coverage in time}
    \label{fig:coverage}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward in time}
    \label{fig:reward}
  \end{subfigure}
  \caption{Coverage and reward in time}
  \label{fig:coverage-reward}
\end{figure}
%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test
